### 2015-02-22 fundamentals/impl.tex

Simulation science - like experimental science - has since the beginning had the
problem of identifying trustworthy results from nonsensical results or noise.
Charles Babbage famously articulated this, ``On two occasions I have been asked,
`Pray, Mr. Babbage, if you put into the machine wrong figures, will the right
answers come out?' ... I am not able rightly to apprehend the kind of confusion
of ideas that could provoke such a question.'' \cite{babbage_passages_2011}.
A simulator must both guard against the \emph{garbage in, garbage out} phenomenon, and also assure correctness.

Multiple strategies, collectively known as \emph{\gls{QA}}, have
been invented over the years to mitigate the structural and algorithmic errors
on the part of simulators themselves. These include \emph{\gls{VV}}
\cite{boehm_software_1989}, \emph{\gls{UQ}}
\cite{sacks_design_1989}, testing, and others. \gls{VV} and \gls{UQ} come from
a distinctly computational science background while testing \emph{et al.} come from software development.

Nuclear engineering software quality is often governed by \gls{NQA1}, an
\gls{ASME} specification
whose latest revision appeared in 2009 \cite{asme_nqa-1a-2009_2009}.
Although \gls{NQA1} is primarily
used for designing reactors, it is general enough to apply to
\Cyclus. \Cyclus has adopted an \emph{agile} development process
\cite{larman_agile_2004},
interpreting \gls{NQA1} in a manner similar to the process adopted by the
\gls{DOE} within \gls{NEAMS} \cite{neams_nuclear_2013} or the
PyNE toolkit \cite{biondo_quality_2014}.

Quality assurance is an on-going process throughout the
entire life of the software. As a simulator where most modeling decisions are made
by third party archetype developers or users, the most important feature of
\gls{QA} is verification. Validation may be impossible and \gls{UQ} is beyond
the scope of the kernel.

Verification may be defined as the question, ``Is \Cyclus being built correctly?''
To answer this question, we turn to the software development of notions of testing,
documentation, version control, style guidelines, and continuous integration.
This suite of process controls supplies mechanism for reliable and reproducible
software. The impetus to implement these controls correctly is particularly strong in a scientific
context because of the emphasis on reproducibility and provenance. These verification techniques are
discussed individually in the sections that follow.

Validation, in contrast,  may be defined as the question,
``Is \Cyclus the correct tool?''
Since \Cyclus is alone in its class as an agent-based fuel cycle simulator, longitudinal
validation is not possible. Nonetheless, code-to-code comparisons with fuel cycle
simulators that use other modeling paradigms are underway, if nascent. However, such
exercises are more likely to bring into relief the differences between the modeling
paradigms than supply substantial \gls{QA} and validation.

Lastly, \gls{UQ} is a process that is used on specific simulation
instantiations to statistically determine quality. \gls{UQ} is a process that applies
to the running of many \Cyclus simulations. The kernel itself, as a generic fuel cycle
simulator, cannot make reliable statements about the uncertainty of a particular fuel
cycle. Any \gls{UQ} results are applicable only to the scenario that is under
analysis by the user. \Cyclus is not a panacea for due diligence on behalf of the fuel
cycle modeler. Thus, \gls{UQ} is envisioned to be beyond the scope of core development
and in the domain of the expert user.

@@

Regression tests ensure significant changes do not
occur over the course of \Cyclus development. Such a change is called a
\emph{regression} because the new version is almost always wrong.
Regression tests are implemented similarly to integration tests.
Nose is used to execute full \Cyclus simulations whose results
are then compared. In this category, however, the comparison is done against
the output of the same input file when run with a previous version of \Cyclus,
typically the last released version.
In some sense, regression tests are `dumb' in that they do
not care about the contents of a simulation being correct, only whether or not
it changed. They pick up glaring time-sensitive mistakes that may
not appear elsewhere, though they do not purport to make claims about the
quality of the physics modeled. Regression tests are primarily implemented
in \Cycamore because its archetypes are sophisticated enough
to make interesting input files. Only a handful ($<10$) of time steps are needed
for regression testing.

@@

Version control preserves the development history, or \emph{provenance}, of
source code files. \Cyclus uses a well-established tool called git
\cite{software_freedom_conservancy_git_2014}
for this purpose. Git is currently the most popular utility in the recent wave
of \emph{distributed version control}. Briefly stated, this means that every
user has a complete local copy of the \emph{repository} (the version control
term for the collection of all possible histories).
One feature that git performs exceptionally well is \emph{branching}.
Branches are distinct pathways in the history that diverge from a mainline source
tree and then may be \emph{merged} back into the mainline. Multiple simultaneous
branches may exist at all points in time. Every change to the code is recorded
in the history, along with metadata such as the author, a timestamp, and an
accompanying message. Thus
it is possible for \Cyclus to accurately replay the entirety of who did what to the
code when.

>>>>

Simulation science - like experimental science - must distinguish trustworthy
results from untrustworthy ones.
Charles Babbage famously articulated this, ``On two occasions I have been asked,
`Pray, Mr. Babbage, if you put into the machine wrong figures, will the right
answers come out?' ... I am not able rightly to apprehend the kind of confusion
of ideas that could provoke such a question.'' \cite{babbage_passages_2011}.
A simulator must assure correctness and guard against the \emph{garbage
in, garbage out} phenomenon.

Multiple strategies, collectively known as \emph{\gls{QA}}, have
been invented to mitigate structural and algorithmic errors
in software. These include \emph{\gls{VV}}
\cite{boehm_software_1989}, \emph{\gls{UQ}}
\cite{sacks_design_1989}, testing, and others.

Nuclear engineering software quality is often governed by \gls{NQA1}, an
\gls{ASME} specification
whose latest revision appeared in 2009 \cite{asme_nqa-1a-2009_2009}.
\Cyclus has adopted an \emph{agile} development process
\cite{larman_agile_2004},
interpreting \gls{NQA1} in a manner similar to the process adopted by the
\gls{DOE} within \gls{NEAMS} \cite{neams_nuclear_2013} or by the PyNE toolkit
\cite{biondo_quality_2014}.

Since \Cyclus modeling decisions are made
by third party archetype developers or users, the most important feature of
\gls{QA} is verification. Validation may be impossible and \gls{UQ} is beyond
the scope of the kernel.

Verification may be defined as the question, ``Is \Cyclus being built correctly?''
To answer this question, \Cyclus relies on software development best practices
such as testing,
documentation, version control, style guidelines, and continuous integration to
ensure reliability and reproducibility. These verification techniques are
discussed individually in the sections that follow.

Validation, in contrast,  may be defined as the question,
``Is \Cyclus the correct tool?''
Since \Cyclus is alone in its class as an agent-based fuel cycle simulator, longitudinal
validation is not possible. Nonetheless, code-to-code comparisons with fuel cycle
simulators that use other modeling paradigms are underway
\cite{huff_extensions_2014}. However, such
exercises are more likely to bring into relief the differences between the modeling
paradigms than supply substantial \gls{QA} and validation.

Lastly, \gls{UQ} is a process that is used on specific simulation
instantiations to statistically determine quality. \gls{UQ} is a process that applies
to the running of many \Cyclus simulations. The kernel itself, as a generic fuel cycle
simulator, cannot make reliable statements about the uncertainty of metrics for a particular fuel
cycle. Any \gls{UQ} results are applicable only to the scenario that is under
analysis by the user. \Cyclus is not a substitute for due diligence on behalf of the fuel
cycle modeler. Thus, \gls{UQ} is envisioned to be beyond the scope of core development
and in the domain of the expert user.

@@

Regression tests ensure unexpected backwards-incompatible changes do not
occur over the course of \Cyclus development. Such a change is called a
\emph{regression} because the new version is almost always wrong.
Regression tests are implemented similarly to integration tests.
Nose is used to execute full \Cyclus simulations whose results
are then compared. In this category, however, the comparison is done against
the output of the same input file when run with a previous version of \Cyclus,
typically the last released version.
In some sense, regression tests are `dumb' in that they do
not care about the contents of a simulation being correct, only whether or not
it changed. They pick up glaring time-sensitive mistakes that may
not appear elsewhere, though they do not purport to make claims about the
quality of the physics modeled. Regression tests are primarily implemented
in \Cycamore because its archetypes are sophisticated enough
to make interesting simulations.

@@

Version control preserves the development history, or \emph{provenance}, of
source code files. \Cyclus uses a well-established \emph{distributed version
control} tool called git \cite{software_freedom_conservancy_git_2014} for this
purpose.  Distributed version control allows every
user to have complete local copy of the \emph{repository} (the version control
term for the collection of all possible histories).
One feature that git performs exceptionally well is \emph{branching}.
Branches are distinct pathways in the history that diverge from a mainline source
tree and then may be \emph{merged} back in. Multiple simultaneous
branches may exist at all points in time. Every change to the code is recorded
in the history, along with metadata such as the author, a timestamp, and an
accompanying message. Thus,
it is possible for \Cyclus to accurately replay the entirety of who did what to the
code when.
